---
title: "w31_olympics"
author: "FG"
date: "27 July 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

35 Years of Olympic Games



```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(countrycode)
library(ggflags)
library(ggthemes)
library(hrbrthemes)


options(scipen = 999)
```

```{r}
tuesdata <- tidytuesdayR::tt_load(2021, week = 31)

olympics <- tuesdata$olympics
regions <- tuesdata$regions
```

```{r}
#glimpse(olympics)
```


```{r}
# View(olympics)
```

```{r}
DataExplorer::profile_missing(olympics)
```


```{r}
glimpse(regions)
```

```{r}
DataExplorer::profile_missing(regions)
```

```{r}
regions%>%filter(is.na(region))
```


```{r}
regions <- regions%>%rename(noc="NOC")


my_olympics <- olympics %>% 
  inner_join(regions,by="noc") %>%
  mutate(medal=replace_na(medal,"Absent")) %>%
  filter(!is.na(region)) %>%
  select(-notes,-name,-team,-event) %>% 
  drop_na() 

# DataExplorer::profile_missing(my_olympics)
```


```{r}
skimr::skim_to_list(my_olympics)
```



Converting "region" vector into country code "iso2c" with {countrycodes} to be able to use {ggflags}

some values were not matched unambiguously:  

- Bolivia
- Individual Olympic Athletes
- Kosovo
- Micronesia


I could do a dataset with only "Individual Olympic Athletes" and eventually use it to add some info in a `geom_`

```{r}
IOA <- my_olympics %>% filter(region %in% "Individual Olympic Athletes" ) %>% count(sex)

IOA
```

After a quick look at the "ambiguous_country_codes", just one is relevant with a "Gold medal", then will see how to use it.

```{r}
ambiguous_country_codes <- c("Boliva","Kosovo","Micronesia")

kos_gold <- my_olympics %>% filter(!region %in% "Individual Olympic Athletes",
                                   region %in% ambiguous_country_codes, 
                                   medal=="Gold") 

kos_gold
```

Update "my_olympics" with `countrycode()`:

```{r}
my_olympics <- my_olympics %>%
  filter(!region %in% "Individual Olympic Athletes",
         !region %in% ambiguous_country_codes) 
```


35 Years: 1896 to 2016

```{r}
my_olympics%>%count(year)
```

Age: 11 to 71 (61 different ages)

```{r}
my_olympics %>% count(age)
```

## Section Age & Sex

```{r}
age_sex_plot <- my_olympics %>% 
  group_by(sex) %>%
  summarize(age,height,weight,year) %>% # count(age)
  ungroup() %>% 
  ggplot(aes(x=factor(age),fill=factor(sex))) + 
  geom_bar(position="stack") +
  scale_fill_fivethirtyeight() +
  labs(title="Distribution of age by sex",
       subtitle="61 different age years form 11 to 71 years old gamers",
       fill="Sex") + 
  theme_fivethirtyeight() +
  theme(axis.text.x = element_text(angle=20),
        legend.position = "bottom",
        plot.title = element_text(size=16,vjust=-0.5),
        panel.grid.major.x = element_blank())

age_sex_plot
```


Composition of plots and background

```{r message=FALSE, warning=FALSE}
library(magick)
library(ggimage)
library(ggpubr)
library(cowplot)
library(extrafont)
library(showtext)
fonts()
olympics_family<-"Roboto Condensed"
#library(ggflags)
library(countrycode)

```



```{r}
olympics_plot <- ggplot() +
  geom_blank() + 
  theme_void() +
  theme()


img_olympics<-"olympics.png"

plot <- ggbackground(age_sex_plot, img_olympics,alpha=.4, color="#CD919E")


plot +
  theme(plot.background = element_rect())#fill="#FFEFDB"))
  
```

Then add a secon plot on the right corner


## Section medals

"At the 1968 Summer Olympics in Mexico City, 29 events in swimming were contested. There was a total of 468 participants from 51 countries competing. The United States dominated the competition, winning 52 of 87 possible medals. 15-year-old phenom Debbie Meyer from Maryland won three gold medals."

source: [Swimming at the 1968 Summer Olympics](https://en.wikipedia.org/wiki/Swimming_at_the_1968_Summer_Olympics#:~:text=At%20the%201968%20Summer%20Olympics%20in%20Mexico%20City%2C%2029%20events,Maryland%20won%20three%20gold%20medals.)


```{r}
my_olympics_with_country_codes<- my_olympics%>%
  mutate(country_code = countrycode(region, 
            origin = 'country.name', 
            destination = 'iso2c'),
         country_code=tolower(country_code)) %>%
  select(year,sport,medal,country_code) %>%
  filter(medal=="Gold") %>%
  count(year,sport,country_code) %>%
  arrange(year) 
  


my_olympics_with_country_codes
```

```{r}
my_favourites <- c("Athletics","Wrestling","Swimming","Shooting","Rowing",
                   "Boxing","Canoeing","Cycling")


my_favourite_sports <- my_olympics_with_country_codes%>%
  filter(sport %in%my_favourites)
```



```{r}
library(ggflags)
```


```{r}
top_golden_sports <- ggplot(data=my_favourite_sports,
       aes(x=factor(year),y=fct_reorder(sport,-year))) +
  geom_point(shape = 21, colour = "gold", fill = NA, size = 7, stroke = 1) +
  ggflags::geom_flag(aes(country=country_code), size=4.5) +
  ggflags::scale_country() +
  guides(country="none") +
  labs(title="Top 8 Gold medal sport winners",
       x="",y="") +
  theme_fivethirtyeight()+
  theme(axis.text.x = element_text(angle=90),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(color="#CD919E"),
        plot.title.position = "plot",
        plot.background = element_blank(),
        panel.background = element_blank())

# top_golden_sports

```


```{r}
#png("top_golden_sports.png")
ggsave("top_golden_sports.png", width = 8, height = 6)

print(top_golden_sports)
dev.off()

```


```{r message=FALSE, warning=FALSE}
library(cowplot)
```



```{r}
img <- "top_golden_sports.png"

inset.plot <- ggdraw() +
  draw_image(img,  x = 0.22, y = -0.2,  scale = .70) 

inset.plot <- inset.plot + theme(legend.position = "none",
                                    plot.background = element_blank(),
                                    panel.background = element_blank())
inset.plot2 <- age_sex_plot + theme(legend.position = "none",
                                    plot.background = element_blank(),
                                    panel.background = element_blank())

double_plot <- ggdraw() +
  draw_plot(inset.plot2, x = 0, y = 0, width = 1, height = 1) +
  draw_plot(inset.plot, x = -0.35, y = .18, width = 1.5, height = 0.9) 


double_plot
```


```{r}
img_olympics<-"olympics.png"
plot <- ggbackground(double_plot, img_olympics,alpha=.2, color="#CD919E")


plot +
  theme(plot.background = element_rect(fill="#FFEFDB"))
```




```{r}
graphics <- ggarrange(plot) 

annotation_plot <- annotate_figure(graphics,
               top = text_grob("Olympics outlook 1896 - 2016  ",color =c("#FF4040", "#FFFFFF", "#FFFFFF"), 
                               face = "bold", size = 40,family=olympics_family),
               bottom = text_grob("DataViz: @fgazzelloni DataSource: \n TidyTuesday week31, Olympic Medals, Kaggle, Financial Times & FiveThirtyEight",
                                  color = c("black"),family=olympics_family,
                                  hjust = 0.5, x = 0.5, face = "bold.italic", size = 10),
               left = text_grob("", color = c("#778899"), rot = 90,size=1),
               right = text_grob(bquote(""), color=c("#778899"),rot = 90,size=1),
               fig.lab = "TidyTuesday week31\n", fig.lab.face = "bold.italic",fig.lab.size=7,
               fig.lab.pos="bottom.right"
)

final_plot <- annotation_plot +
  annotate(geom = "text", label="The historical dataset on the modern Olympic Games,\n including all the Games from Athens 1896 to Rio 2016",x = 0.7, y = 0.81, 
           colour = "black", size = 6,family=olympics_family) 
  

final_plot
```

Attach the Olympic logo at the sides of the title

```{r}
img_olympics<-"olympics.png"

final <- ggdraw() +
  draw_image(img_olympics,  x = -0.35, y = 0.45, scale = .10) +
  draw_image(img_olympics,  x = 0.32, y = 0.45, scale = .10) +
  draw_plot(final_plot)


```


Saving:
```{r}
ragg::agg_png(here::here("w31", "w31_olympics.png"),
              res = 320, width = 14, height = 8, units = "in")
final

dev.off()

```


