---
title: "#TidyTuesday Week 15 US Egg Production & #30daychartchallenge day21 down/upwards"
author: "Federica Gazzelloni"
execute:  
  comments: ""
  eval: true
  echo: true
  warning: false
  message: false
---


> To cite Federica's work, please use: Gazzelloni F., 2023 Data Visualization US Egg Production down/upwards

```{r}
library(tidyverse)
tuesdata <- tidytuesdayR::tt_load(2023, week = 15)
```

```{r}
eggs <- tuesdata$`egg-production`
cages <- tuesdata$`cage-free-percentages`
```

```{r}
eggs%>%
  inner_join(cages,by="observed_month")%>%
  distinct() %>%
  mutate(ymon=zoo::as.yearmon(observed_month),
         .after=observed_month)
```


```{r}
df <- eggs%>%
  inner_join(cages,by="observed_month")%>%
  distinct() %>%
  mutate(ymon=zoo::as.yearmon(observed_month),
         .after=observed_month)%>%
  select(-source.x,-source.y,-observed_month)%>%
  group_by(ymon,prod_process)%>%
  mutate(pct_hens=round(n_hens/sum(n_hens)*100,2),
          pct_eggs=round(n_eggs/sum(n_eggs)*100,2))%>%
  arrange(ymon)

df%>%head
```

```{r}
df %>%
  ggplot(aes(ymon,n_hens))+
  geom_line(aes(group=prod_type))+
  facet_wrap(~prod_process,scales = "free")
```

```{r}
df %>%
  filter(!prod_process=="all")%>%
  ggplot()+
  geom_line(aes(ymon,n_hens,group=prod_type),
            lineend = "butt",
            arrow = arrow(length = unit(0.05,"inches"),
                          type="closed"))+  
  facet_wrap(~prod_process,scales = "free_y")+
  scale_y_log10()
```

```{r}
df %>%
  filter(!prod_process=="all")%>%
  ggplot(aes(ymon,n_hens))+
  geom_line(aes(group=prod_type))+
  facet_wrap(~prod_process,scales = "free_y")
```

```{r}
df %>%
  filter(!prod_process=="all")%>%
  ggplot()+
  geom_line(aes(ymon,percent_eggs,group=prod_type),
            lineend = "butt",
            arrow = arrow(length = unit(0.05,"inches"),
                          type="closed"))+  
  geom_line(aes(ymon,percent_hens,group=prod_type),
            lineend = "butt",
            arrow = arrow(length = unit(0.05,"inches"),
                          type="closed"))+
  facet_wrap(~prod_process,scales = "free_y")
```



### Fill missing values

you can fill in the data with two approaches:

- Up - While filling the missing values, you have to specify the direction of filling of values. If you choose Up, then the filling process will be bottom-up.
- Down - In this method, you have to set the direction of filling to down.

source: [https://www.digitalocean.com/community/tutorials/fill-missing-values-in-tidyr-r](https://www.digitalocean.com/community/tutorials/fill-missing-values-in-tidyr-r)


Use `.direction = "downup"` to fill missing values in both directions


```{r}
df %>%
  filter(!prod_process=="all")%>%
  tidyr::fill(percent_eggs, .direction = 'downup') %>% # View()
  mutate(percent_eggs=ifelse(is.na(percent_eggs),25,percent_eggs))%>% #View
  ggplot()+
  geom_line(aes(ymon,percent_eggs,group=prod_type),
            color="red",
            lineend = "butt",
            arrow = arrow(length = unit(0.05,"inches"),
                          type="closed"))+  
  geom_line(aes(ymon,percent_hens,group=prod_type),
            lineend = "butt",
            arrow = arrow(length = unit(0.05,"inches"),
                          type="closed"))+
  facet_wrap(~prod_process,scales = "free_y")+
  ggthemes::theme_fivethirtyeight()
```
Aug 2016 - Feb 2021
```{r}
df %>%
  filter(!prod_process=="all")%>% # count(ymon)
  tidyr::fill(percent_eggs, .direction = 'downup') %>% # View()
  mutate(percent_eggs=ifelse(is.na(percent_eggs),25,percent_eggs))%>% #View
  pivot_longer(cols = c("percent_eggs","percent_hens"))%>%
  mutate(name=ifelse(name=="percent_eggs","eggs (%)","hens (%)")) %>%
  ggplot()+
  geom_line(aes(ymon,value,group=name,color=name),
            lineend = "butt",
            arrow = arrow(length = unit(0.05,"inches"),
                          type="closed"))+ 
  facet_wrap(~prod_process,scales = "free_y")+
  labs(title="US Egg Production Aug 2016 - Feb 2021",
       subtitle="Hens & Eggs down/upwards",
       color="",
       caption="DataSource: #TidyTuesday Week 15 US Egg Production\n #30DayChartChallenge Day 21 down/upwards | DataViz: Federica Gazzelloni")+
  ggthemes::scale_color_fivethirtyeight()+
  ggthemes::theme_fivethirtyeight()+
  theme(strip.background = element_rect(color="#192d40",fill="#192d40"),
        strip.text = element_text(color="white"))
```

```{r}
ggsave("w15_UEPD.png",width = 7,height = 5,bg="white")
```

