---
title: "w22 Company reputation poll"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Axios and Harris Poll to gauge the reputation of the most visible brands in America, based on 20 years of Harris Poll research. 
Based on a survey of 33,096 Americans in a nationally representative sample conducted March 11-April 3, 2022. 

Each company received approximately 325 ratings. An RQ score is calculated by: [ (Sum of ratings of each of the 9 attributes)/(the total number of attributes answered x 7) ] x 100.

Score ranges: 80 & above: Excellent | 75-79: Very Good | 70-74: Good | 65-69: Fair | 55-64: Poor | 50-54: Very Poor | Below 50: Critical
```{r}
library(tidyverse)
```

```{r}
poll <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-05-31/poll.csv')
reputation <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-05-31/reputation.csv')
```
```{r}
poll%>%head
```
```{r}
reputation%>%head
```
```{r}
library(extrafont)
# loadfonts()
```

```{r}
p<-reputation%>%
  arrange(rank) %>%
  filter(!industry=="Other")%>%
  mutate(name=reorder(name,score),
         industry=factor(industry),
         industry=reorder(industry,-score))%>%
  ggplot(aes(x=score,y=rank,color=name))+
  geom_point(size=1,shape=21,stroke=0.5)+
  geom_line(linetype="dashed")+
  scale_y_reverse(limits=c(100,1))+
  facet_wrap(~industry, ncol = 6)+
  ggthemes::scale_color_colorblind()+
  guides(color=guide_legend(nrow=1,byrow=TRUE,title.position = "left"))+
  coord_polar(theta = "y",start = 1000, direction = 1, clip = "off")+
labs(title="Corporate Reputation Rankings - 2022",
     subtitle="100 most visible companies grouped by 18 industries are scored in 7 categories",
     caption="\nDataSource: #TidyTuesday 2022 week22 - Axios and Harris Poll\nDataViz: Federica Gazzelloni (@fgazzelloni)",
     x="",
     y="Rank vs Score polarized dotted lines shows Citizenship is the dominant reputation category for all industries, followed by Ethics and Trust,\nCulture, P&S(Product and Service), Growth, and Vision, in this order for most of the industries.\nIndustries are ordered by the highest score received by each reputation category.\nRetail shows a full score trend contrasting Insurance with just one rank/score for each category.",
     color="Reputation category")+
  theme_linedraw()+
  theme(text=element_text(family="Roboto Condensed"),
        axis.title = element_text(size=10,hjust=0),
        plot.title = element_text(size=30),
        plot.subtitle = element_text(size=14),
        legend.position = "top",
        panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        strip.background = element_rect(color="grey20",fill="grey20"),
        plot.background = element_rect(color="grey30"))
```




```{r}
ggsave("w22_reputation2.png",
       dpi=320,
       width = 9,
       height = 7)
```

