---
title: "Week 52 Star Trek Timelines"
author: "fg"
date: "2023-04-02"
output: html_document
---


## #30DayChartChallenge - Day2 - Waffle
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = "",
                      cache = TRUE)
```

Load the libraries
```{r eval=TRUE}
library(tidyverse)
```



```{r eval=TRUE}
tuesdata <- tidytuesdayR::tt_load(2022, week = 52)
tlBooks <- tuesdata$tlBooks
tlFootnotes<-tuesdata$tlFootnotes
```

```{r eval=TRUE}
tlBooks%>%names
```

```{r eval=TRUE}
tlFootnotes%>%names
```

Join the two sets by `footnote`

The new dataset combines, year, title, ... with the footnote of the `Star Trek Timelines`.
The data comes from the {rtrek} package by Georgios Karamanis.

```{r eval=TRUE}
df <- tlBooks %>%
  inner_join(tlFootnotes,by="footnote")

df%>%DataExplorer::profile_missing()%>%arrange(pct_missing)
```

## How to make a `Waffle`

This is a little example from: [https://r-charts.com/part-whole/waffle-chart-ggplot2/](https://r-charts.com/part-whole/waffle-chart-ggplot2/)

```{r eval=TRUE}
# install.packages("waffle", repos = "https://cinc.rud.is")
library(waffle)

# Vector
x <- c(30, 25, 20, 5)

# Waffle chart
waffle(x, rows = 8)
```


In this dataset there are three formats: book, episode and story
```{r}
df%>%
  count(format)%>%
  waffle(rows=20)
```

### Using `ggplot2`

This `Waffle` is made of 12 different colors for identifying the `SSeries`. Here are used many colors from the `trekcolors` package for coloring the `series` of different colors. 

```{r eval=TRUE}
# install.packages("trekcolors")
library(trekcolors)
# trekcolors::lcars_colors()
```


The fonts are from the `trekfont` package.

```{r eval=TRUE}
# install.packages("trekfont")
library(trekfont)
# trekfont::show_trekfonts()
library(showtext)
font <- c("Khan", "StarNext", "FederationDS9Title", "Federation", "Klingon", "ModernVulcan", "TNGcast", "FederationStarfleet")
path <- system.file(paste0("fonts/", font, ".ttf"), package = "trekfont")
for(i in seq_along(font)) font_add(font[i], path[i])
font_families()
showtext_auto(enable = TRUE)
```



```{r eval=TRUE}
library(waffle)
df%>%
  count(series)%>%
  drop_na()%>%
  waffle(rows = 20, size = 0.5)+
  scale_fill_manual(values =as.character(lcars_colors())) +
  # Waffle plot
  #ggplot(aes(fill = series, values = n)) +
  #geom_waffle(n_rows = 20, size = 0.5, colour = "white") +
  #scale_fill_manual(values =as.character(lcars_colors())) +
  coord_equal() +
  scale_x_continuous(expand = c(0, 0))+
  labs(title="Star Trek Timelines Series",
       subtitle = "",
       caption="DataSource: #TidyTuesday 2022 week52 - Star Trek Timelines\nDataViz: Federica Gazzelloni #30DayChartChallenge 2023 Day2 - Waffle\n")+
  theme_void()+
  theme(text = element_text(family= "StarNext",size=14),
        #legend.position = "bottom",
        plot.title = element_text(size=50,hjust = 0.3,vjust = 0),
        plot.caption = element_text(size=20,hjust = 0.4,family="FederationDS9Title"),
        panel.background = element_rect(fill="#9977AA",color="#9977AA"),
        plot.background = element_rect(fill="#9977AA",color="#9977AA"))
```

```{r}
ggsave("ss.png",
       width = 6,height = 5.5,
       bg="#9977AA",
       dpi=200)
```

