---
title: "drought"
author: "FG"
date: "7/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

Resources for this week:

- [rfordatascience](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-07-20/readme.md)
- [ComprehensiveStatistics](https://droughtmonitor.unl.edu/DmData/DataDownload/ComprehensiveStatistics.aspx)
- [NYT](https://www.nytimes.com/interactive/2021/06/11/climate/california-western-drought-map.html)
- [CNN](https://edition.cnn.com/2021/06/17/weather/west-california-drought-maps/index.html)
- [Drought Classification](https://droughtmonitor.unl.edu/About/AbouttheData/DroughtClassification.aspx)
- [drought Monitor](https://droughtmonitor.unl.edu/ConditionsOutlooks/CurrentConditions.aspx)

Facts 2001 to 2021:

- Six states are now entirely in drought conditions
- The drought has nearly doubled in size from this time last year
- Around 25% of the country was in drought conditions in July 2020
- More than 94% of the West is in drought
- More than 60% of the region is in 'extreme' or 'exceptional' drought
- Six states completely in drought conditions; California, Oregon, Nevada, Utah, Idaho and North Dakota

Data limitations:

- statistics are limited to areas based on counties
- population changes over time
- population is distributed evenly across each county

- [limitation](https://droughtmonitor.unl.edu/About/AbouttheData/PopulationStatistics.aspx)

Caption: The U.S. Drought Monitor is jointly produced by the National Drought Mitigation Center at the University of Nebraska-Lincoln, the United States Department of Agriculture, and the National Oceanic and Atmospheric Administration. Map courtesy of NDMC.

```{r}
tuesdata <- tidytuesdayR::tt_load(2021, week = 30)

drought <- tuesdata$drought
```

```{r}
glimpse(drought)
```

```{r}
drought%>%count(drought_lvl)
```

Intensity and Impacts
```{r}
intensity_impacts<- c("D0"="Abnormally Dry",
                      "D1"="Moderate Drought",
                      "D2"="Severe Drought",
                      "D3"= "Extreme Drought",
                      "D4"="Exceptional Drought")
```


```{r}
DataExplorer::profile_missing(drought)
```

```{r}
drought_short <- drought %>%
  dplyr::select(-map_date,-stat_fmt) %>%
  filter(!drought_lvl=="None",!area_pct==0)

drought_short
```


```{r message=FALSE, warning=FALSE}
library(sf)
library(raster)
library(spData)
library(spDataLarge)

library(maps)
library(viridis)
library(ggthemes)
```

```{r}
us_county_map <- map_data("county")

county_plot<-ggplot()+
  geom_polygon(data=us_county_map,aes(x=long,y=lat,group = group),
               fill=NA,color = "lightblue") +
  theme_map()

county_plot
```


```{r}
my_states <- drought_short%>%count(state_abb)
```



```{r}
us_state_map <- map_data("state")

state_plot<-ggplot()+
  geom_polygon(data=us_state_map,aes(x=long,y=lat,group = group),
               fill=NA,color = "lightblue") +
  theme_map()

state_plot
```



```{r}
us_plot<-ggplot()+
  geom_polygon(data=us_county_map,aes(x=long,y=lat,group = group),
               fill=NA,color = "lightblue") +
   geom_polygon(data=us_state_map,aes(x=long,y=lat,group = group),
               fill=NA,color = "pink") +
  theme_map()

us_plot
```



```{r}
library(zipcodeR)
zipcodeR::download_zip_data()

geo_codes<- zipcodeR::search_state(drought_short$state_abb)%>%
  dplyr::select(major_city,county,state,lat,lng,
         population,population_density,
         land_area_in_sqmi,water_area_in_sqmi,
         housing_units,occupied_housing_units,
         median_home_value,median_household_income) %>%
  drop_na()

geo_codes
```



```{r}
my_geo_codes_df<-geo_codes%>%
  dplyr::select(state,lat,lng)
  

drought_short_map <- drought_short %>%
  arrange(valid_start)%>%
  mutate(year=lubridate::year(valid_start),
         month=lubridate::month(valid_start))%>%
  filter(str_detect(valid_start,"2021")) %>%
  #filter(month==c(1,2,3)) %>%
  group_by(month,state_abb,drought_lvl) %>%
  summarize(med_area_pct=round(median(area_pct),2))%>%
  ungroup() %>%
  filter(!med_area_pct==0) %>%
  left_join(my_geo_codes_df,by=c("state_abb"="state")) %>%
  mutate(month = month.name[month])
  
drought_short_map
```



```{r}
library(extrafont)
loadfonts()
fonts()

droughts_family <- "Roboto Condensed"
```


```{r}
drought_plot <- ggplot() +
  geom_point(data=subset(drought_short_map,lat>25&lat<50),
                         aes(x=lng,y=lat,color=drought_lvl),
             alpha=0.5,size=.4) +
  geom_polygon(data=us_county_map,aes(x=long,y=lat,group = group),
               fill=NA,color = "lightblue",size=0.2) +
   geom_polygon(data=us_state_map,aes(x=long,y=lat,group = group),
               fill=NA,color = "pink",size=0.4) +
  facet_wrap(~factor(month, levels=c('January','February','March','April',
                                     'May','June','July'))) +
  labs(title="",
       subtitle="",
       caption="US Droughts map: available values by County\n Jan to July 2021",
       #tag = "Jan to July 2021",
       color="Level") +
  scale_color_viridis(labels = intensity_impacts,discrete = TRUE) +
  guides(color = guide_legend(override.aes = list(size = 3))) +
  ggthemes::theme_map() +
  theme(legend.position = "top",
        legend.title = element_text(family = droughts_family),
        legend.text = element_text(size=8,family =droughts_family),
        legend.background = element_blank(),
        legend.box.background = element_blank(),
        legend.key = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(family = droughts_family),
        plot.title =element_text(size=15,face="bold",family =droughts_family,color="black"),
        plot.subtitle =element_text(size=12,face="bold",family =droughts_family),
        plot.caption =element_text(size=9,family =droughts_family,hjust = 0),
        #plot.tag = element_text(size=9,face="bold",family =droughts_family,hjust = 0),
        plot.caption.position = "panel",
        plot.title.position = "panel")
  


# drought_plot 
```



```{r message=FALSE, warning=FALSE}
library(tidymodels)
tidymodels_prefer()


cooked_drought <- recipe(drought_lvl ~ med_area_pct+month + lat+lng +state_abb,drought_short_map) %>%
  prep()%>%
  bake(new_data=NULL)

cooked_drought
```


```{r}
cooked_drought_plot <- cooked_drought%>%
  group_by(drought_lvl,month) %>%
  summarise(med_med_area_pct=median(med_area_pct)) %>%
  mutate(month_f= case_when(month=="January"~1,
                            month=="February"~2,
                            month=="March"~3,
                            month=="April"~4,
                            month=="May"~5,
                            month=="June"~6,
                            month=="July"~7)) %>%

  ggplot(aes(x=month_f,y=med_med_area_pct,fill=month)) +
  geom_col()+
  geom_text(aes(label=month),nudge_y = 1.5,size=1.5,family = droughts_family) +
  geom_text(aes(label=med_med_area_pct),nudge_y = -1.5,size=1.5,color="white") +
  facet_wrap(~factor(drought_lvl,labels = intensity_impacts),
             nrow = 1,ncol = 5,strip.position="bottom",
             scales = "fixed")+
  scale_fill_viridis(discrete = TRUE) +
  labs(y="",
       caption="Forecasted Droughts affected area pct median value by first 7 months of the year")+
  theme_void()+
  theme(legend.position = "none",
        plot.caption = element_text(family = droughts_family),
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        strip.placement = "inside",
        strip.text = element_text(family = droughts_family,size=6),
        plot.background = element_blank())


  
# cooked_drought_plot
```


```{r}
library(cowplot)
```


```{r}
inset.plot <- cooked_drought_plot + theme(legend.position = "none")

double_plot <- ggdraw() +
  draw_plot(drought_plot,width = 1, height = 1) +
  draw_plot(inset.plot, x = 0.39, y = .035, width = .55, height = .25)

# double_plot
```

```{r}
library(ggpubr)
library(ggimage)
```


```{r}
graphics <- ggarrange(double_plot) 

annotation_plot <- annotate_figure(graphics,
               top = text_grob("US droughts monitor condition outlook: ",color =c("#36648B", "#607B8B", "#668B8B"), 
                               face = "bold", size = 30,family=droughts_family),
               bottom = text_grob("DataViz: @fgazzelloni DataSource: \n TidyTuesday week30, US Droughts,Drought Monitor,NYTimes & CNN",
                                  color = c("#36648B", "#607B8B", "#668B8B"),family=droughts_family,
                                  hjust = 0.5, x = 0.5, face = "bold.italic", size = 10),
               left = text_grob("", color = c("#778899"), rot = 90,size=1),
               right = text_grob(bquote(""), color=c("#778899"),rot = 90,size=1),
               fig.lab = "TidyTuesday week30\n", fig.lab.face = "bold.italic",fig.lab.size=7,
               fig.lab.pos="bottom.right"
)

final_plot <- annotation_plot +
  annotate(geom = "text", label="map approximates drought-related impacts",x = 0.7, y = 0.87, 
           colour = "#BF3EFF", size = 8,family=droughts_family) 
  

# final_plot
```


```{r}
img <- png::readPNG('NDMC-logo-usdm-opt.png')
img1 <- png::readPNG('DOC-logo-usdm.png')
img2 <- png::readPNG('NOAA-logo-usdm.png')
img3 <- png::readPNG('USDA-logo-usdm-opt.png')
```



```{r width=100}
final <- ggdraw() +
  draw_image(img,  x = -0.45, y = 0.44, scale = .10) +
  draw_image(img1,  x = -0.38, y = 0.44, scale = .10) +
  draw_image(img2,  x = 0.33, y = 0.44, scale = .10) +
  draw_image(img3,  x = 0.42, y = 0.44, scale = .10) +
  draw_plot(final_plot)

# final
```



Saving:
```{r}
ragg::agg_png(here::here("w30", "w30_drought.png"),
              res = 320, width = 14, height = 8, units = "in")
final

dev.off()

```






